---
title: "Catch comparison with subsampling"
author: "Mollie Brooks"
date: "30 Apr 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{catchcomparison}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

#Preliminaries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

```{r libs}
library(selfisher)
library(plyr)
library(ggplot2); theme_set(theme_bw())
library(bbmle)
library(knitr)
library(splines) 
```

Read in data that is distributed with the package.
```{r data, echo=TRUE}
data(ccmhsampdat)
dat=ccmhsampdat
head(dat)
```

##Data transformations
The length predictor might need to be scaled for numerical stability.
```{r scale}

mu=mean(rep(dat$length, dat$total))
var=var(rep(dat$length, dat$total))
  
dat=transform(dat, 
				q=sampling_test1/sampling_test2,
				sl=(length-mu)/sqrt(var))
```

For plotting, we need to aggregate the hauls and use the sampling fractions to raise the counts.
```{r aggregate}
sumdat=ddply(dat, ~length+sl, summarize, 
				prop=sum(test1)/sum(total), 
				total=sum(total),
				raised_prop=sum(test1/sampling_test1)/
						(sum(test1/sampling_test1)+sum(test2/sampling_test2)),
				raised_total=sum(test1/sampling_test1)+sum(test2/sampling_test2),
				raised_ratio = sum(test1/sampling_test1)/sum(test2/sampling_test2))
```

#Model
```{r fits, warning=FALSE}
m1=selfisher(prop~offset(log(q))+sl+I(sl^2)+I(sl^3)+I(sl^4), total=total, dat, haul=haul)
```

#Predictions
Make predictions of the catch ratio (test1/test2) on a new dataset as if the sampling ratio (i.e. `q`) was 1.

```{r pred}
newdata=data.frame(length=unique(dat$length))
newdata=transform(newdata,
				sl=(length-mu)/sqrt(var),
				total=1,
				haul=1,
				q=1
)
newdata$raised_ratio=predict(m1, newdata=newdata, type="ratio")
```

#Plot predictions
```{r plot}
p1=ggplot(sumdat, aes(length, raised_ratio))+
	geom_point(aes(size=total))+
   geom_line(data=newdata)+
	geom_hline(yintercept = 1, lty=2)
p1
```


#Bootstrap to get CI on predictions
```{r ci}
bs=bootSel(m1, nsim=1000, parallel = "multicore", ncpus = 4, FUN=function(mod){predict(mod, newdata=newdata, type="ratio")})

quants=apply(bs$t, 2, quantile, c(0.025, 0.5, 0.975))
newdata[,c("lo", "mid", "hi")]=t(quants)

ggplot(sumdat, aes(length, raised_ratio))+
	geom_point(aes(size=total))+
   geom_line(data=newdata)+
   geom_ribbon(data=newdata, aes(ymin=lo, ymax=hi), alpha=0.2)+
	geom_hline(yintercept = 1, lty=2)
```

#Splines instead of polynomial

#Models
```{r spline_fits}
s3=selfisher(prop~offset(log(q))+ns(sl, 3), psplit=FALSE, total=total, dat, haul=haul)
s4=selfisher(prop~offset(log(q))+ns(sl, 4), psplit=FALSE, total=total, dat, haul=haul)
s5=selfisher(prop~offset(log(q))+ns(sl, 5), psplit=FALSE, total=total, dat, haul=haul)
AICtab(s3, s4, s5, m1)
```

#Predictions
```{r spline_pred}
newdata=data.frame(length=unique(dat$length))
newdata=transform(newdata,
          sl=(length-mean(dat$length))/sqrt(var(dat$length)),
          total=1,
          haul=1,
					q=1
)
newdata$raised_ratio=predict(s3, newdata=newdata, type="ratio")
```

#Bootstrap to get CI on predictions
In this vignette, we onlt do 100 simulations, but really there should be 1000.
```{r spline_ci}
bs=bootSel(s3, nsim=100, parallel = "multicore", ncpus = 4, FUN=function(mod){predict(mod, newdata=newdata, type="ratio")})

quants=apply(bs$t, 2, quantile, c(0.025, 0.5, 0.975))
newdata[,c("lo", "mid", "hi")]=t(quants)
```

#Plot predictions
```{r spline_plot}
p2=ggplot(sumdat, aes(length, raised_ratio))+
	geom_point(aes(size=total))+
   geom_line(data=newdata)+
   geom_ribbon(data=newdata, aes(ymin=lo, ymax=hi), alpha=0.2)+
	geom_hline(yintercept = 1, lty=2)
p2
```
