---
title: "Troubleshooting with selfisher"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{troubleshooting}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo=FALSE}
library(selfisher)
```

This vignette covers common problems that occur while using `selfisher`. 
The contents will expand with experience.

#Warnings

##Model convergence problem; non-positive-definite Hessian matrix

This often occurs when a model is overparameterized (i.e. the data does not contain information to estimate the parameters).
Plotting the data against potential covariates should help to avoid fitting unreasonable models that don't converge.

Scaling predictor variables may help.
```{r scale}
m0=selfisher(nwide/(nfine+nwide)~scale(Lengths), total=nfine+nwide, haddock)
```


Models with non-positive definite Hessian matricies should be excluded from further consideration, in general.  

##Model convergence problem; Eigen value problems

Models with convergence problems should be excluded from further consideration, in general.

##NA/NaN function evaluation
```{r NA function, eval=FALSE}
Warning in nlminb(start = par, objective = fn, gradient = gr) :
  NA/NaN function evaluation
```
This warning often occurs when the optimizer wanders into a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above. 

The following warnings can be treated in the same way as an NA/NaN function evaluation:
```{r Cholmod, eval=FALSE}
Cholmod warning 'matrix not positive definite'
```
```{r lgamma, eval=FALSE}
Warning in f(par, order = order, ...) : value out of range in 'lgamma'
```
